---
title: "Incorporating lipoprotein(a) into cardiovascular risk prediction equations"
subtitle: "Product: Olpasiran; Protocol Number: 20250105"
author: "[Byron C. Jaeger, PhD](https://www.byronjaeger.com/)"
date: "Document created on `r format(Sys.time(), '%B %d, %Y')`" 
bibliography: refs.bib
csl: refs.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_arial.docx
    keep_md: true
    toc: true
    toc_depth: 2
---

```{r setup-knitr, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,    # don't show code blocks
  message = FALSE, # don't print messages to the doc
  warning = FALSE, # don't print warnings to the doc
  dpi = 300,       # higher dpi => higher render time 
  fig.width = 6,   # default figure width in inches
  fig.height = 4   # default figure height in inches
)

# Hello! This is the template word document for Perisphere reports.
# There are several code chunks used to set up the format of the
# document below. You don't have to modify those unless you want 
# to change the format of your output. 

```

```{r setup-pkgs}

# packages are usually handled by packages.R (in main directory).
# When tar_make() is run, it loads everything in packages.R.
# Assuming this document is in the list of targets (see ?tar_render),
# you won't need to include the library calls below.
# however, the template includes packages in its initial form
# so that you can knit the document without having to run tar_make().
# The library calls can safely be removed once the required packages
# are in packages.R and this doc is part of the targets pipeline.

# standards for word docs
library(officer)
library(flextable)

# specify a rounding specification 
library(table.glue)

# general purpose
library(tidyverse)

# table 1
library(gtsummary)

```

```{r setup-rounding}


rspec_4d <- round_spec() %>% 
  round_using_decimal(digits = 4)

rspec_3d <- round_spec() %>% 
  round_using_decimal(digits = 3)

rspec_2d <- round_spec() %>% 
  round_using_decimal(digits = 2)

```


```{r setup-figures}

set_fig_landscape <- function() {
  knitr::opts_chunk$set(fig.width = 11, fig.height = 6)
}

set_fig_portrait <- function() {
  knitr::opts_chunk$set(fig.width = 6, fig.height = 4)
}

theme_set(
  new = theme_bw() + 
    theme(panel.grid = element_blank(),
          text = element_text(size = 12))
)

```

```{r setup-margins}

# Set a default footer because when we use block_section, it 
# starts a new style template for headers and footers. Thus,
# if we want page numbers to appear on every page, we need to
# put that into the default footer and then put that default
# footer into every block section. MS word...¯\_(ツ)_/¯
# note to future self - go here if you want to remember this: 
# https://github.com/davidgohel/officedown/discussions/103

footer_default <- block_list(
  fpar(
    run_word_field(field = "PAGE"),
    fp_p = fp_par(text.align = "left") 
  )
)

# Code in this chunk sets the margins for your document. 
# The helper functions page_long_above() and page_wide_above()
# also provide shortcuts for switching between wide (landscape)
# and normal (long) MS word layouts. 

# helper function for defining margins
make_margin <- function(x=NULL){
  
  if(is.null(x)){ return(page_mar()) } 
  
  if(is.character(x)){
    
    out <- switch(
      tolower(x),
      "normal" = page_mar(1, 1, 1, 1),
      "narrow" = page_mar(1/2, 1/2, 1/2, 1/2),
      "moderate" = page_mar(1, 1, 3/4, 3/4)
    )
    
  }
  
}

# set content above to be portrait
page_long_above <- function(margins = make_margin('narrow')){
  
  # b/c block_sections apply to content ABOVE, we set
  # the content BELOW to be landscape when we use page_long_above  
  set_fig_landscape()
  
  block_section(
    prop_section(
      page_size = page_size(orient = "portrait"),
      type = "continuous",
      page_margins = margins,
      footer_default = footer_default
    )
  )
  
}

# set content above to be landscape
page_wide_above <- function(margins = make_margin('narrow')){
  
  # b/c block_sections apply to content ABOVE, we set
  # the content BELOW to be portrait when we use page_wide_above
  set_fig_portrait()
  
  block_section(
    prop_section(
      page_size = page_size(orient = "landscape"),
      type = "continuous",
      page_margins = margins,
      footer_default = footer_default
    )
  )
  
}

```

```{r setup-targets}

# # uncomment when there are targets to load
withr::with_dir(
  new = here::here(),
  code = {
    targets::tar_load(names = c(data_excluded_10yr_risk,
                                data_excluded_30yr_risk,
                                data_train, 
                                data_test,
                                fig_lpa_spline,
                                meta,
                                analyses,
                                evaluations))
    
    exclusions_30yr <- read_csv("data/exclusions_30yr.csv")
    exclusions_10yr <- read_csv("data/exclusions_10yr.csv")
    
  }
)

```


```{r setup-meta}

set_default_dictionary(meta)

```


```{r setup-formatters}

# make a short name for table.glue::table_value - this makes
# the inline reporting of results less complicated to read.
tv <- table.glue::table_value
tg <- table.glue::table_glue

r_version <- tg("{version$major}.{version$minor}")

```

```{r setup-tables}

# Footnotes should be wrapped in `as_paragraph`:

ckd_defn <- as_paragraph("Chronic Kidney Disease was defined by an estimated glomerular filtration rate <60 ml/min/1.73 m", as_sup(2L), " based on the 2021 CKD-EPI creatinine equation.")

# abbreviation management

abbrvs <- list(
  "SD"   = 'standard deviation',
  "ACR" =	"Albumin-to-creatinine ratio",
  "ASCVD" =	"Atherosclerotic cardiovascular disease",
  "CI" =	"Confidence interval",
  "eGFR" =	"Estimated glomerular filtration rate",
  "HbA1c" =	"Glycated hemoglobin",
  "HDL" =	"High-density lipoprotein",
  "LDL" =	"Low-density lipoprotein",
  "SBP" = "Systolic blood pressure",
  "DBP" = "Diastolic blood pressure"
)



abbrvs_paste <- function(strings){
  
  front <- 'Abbreviations'
  
  matter <- glue::glue_collapse(strings, sep = '; ', last = '; and ')
  
  paste(front, matter, sep = ': ')
  
}

abbrvs_write <- function(abbr){
  
  sorted <- abbr[sort(names(abbr))]
  
  strings <- map2_chr(names(sorted), sorted, paste, sep = ' = ')
  
  as_paragraph(abbrvs_paste(strings))
  
}


```

```{r setup-footers}

footer_nri <- as_paragraph(
  paste(
    "NRI⁺ reflects how often participants who experienced an event were",
    "correctly reclassified to a higher predicted risk (minus those moved",
    "lower) by the new model. NRI⁻ reflects how often participants without",
    "events were correctly reclassified to a lower predicted risk (minus those",
    "moved higher). The overall NRI is the sum of NRI⁺ and NRI⁻, summarizing",
    "total improvement in risk classification compared with the reference model."
  )
)

```



```{r create-tbl_characteristics}

tbl_characteristics <- data_train %>% 
  bind_rows(data_test) %>% 
  mutate(risk_horizon = paste("year", risk_horizon, sep = '_')) %>% 
  split(.$risk_horizon) %>% 
  map(translate_data, units = 'descriptive') %>% 
  map(.f =  ~ {
    tbl_summary(.x, 
                by = 'study', 
                include = c("age",
                            "sex",
                            "race",
                            "smoke_current",
                            "bmi",
                            "sbp",
                            "med_htn",
                            "diabetes",
                            "chol_total",
                            "chol_hdl",
                            "med_statin",
                            "egfr_lt60",
                            "lpa_nmol",
                            "lpa_mgdl")) %>% 
      as_flex_table() %>% 
      flextable_polish() %>% 
      flextable_autofit()
  })

```

```{r create-tbl_lpa_lm}

tbl_lpa_lm <- analyses %>% 
  select(.risk_horizon, .lpa_defn, .lpa_trafo, starts_with("fit_lm")) %>% 
  # remove duplicates
  group_by(.risk_horizon, .lpa_defn, .lpa_trafo) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(
    tbl = pmap(
      .l = list(.lpa_defn, .lpa_trafo, fit_lm, fit_lm_tidy),
      .f = function(.lpa_defn, .lpa_trafo, fit_lm, fit_lm_tidy){
        
        r_squared <- table_value(summary(fit_lm)$r.squared * 100)
        
        if(.lpa_trafo == 'log_trafo'){
          lpa_var <- paste(.lpa_defn, "log", sep = '_')
        } else {
          lpa_var <- .lpa_defn
        }
        
        lpa_lab_bare <- meta$get_label(lpa_var)
        lpa_lab_unit <- meta$get_label(lpa_var, units = 'descriptive')
        
        fit_lm_tidy %>% 
          filter(name != '(Intercept)',
                 level != "no" | is.na(level)) %>% 
          select(-level, -term, -std.error, -statistic) %>% 
          mutate(name = translate_names(name, units = 'model'),
                 label = if_else(is.na(label) | label == 'yes', name, label),
                 name = if_else(name == label, NA_character_, name),
                 tbv_pval = table_pvalue(p.value),
                 tbv_est = table_glue("{estimate}\n({conf.low}, {conf.high})",
                                      rspec = rspec_2d),
                 tbv_est = if_else(reference, true = "0 (ref)", false = tbv_est),
                 .keep = 'unused') %>% 
          relocate(tbv_pval, .after = tbv_est) %>% 
          as_grouped_data(groups = 'name') %>% 
          drop_empty_rows() %>%  
          as_flextable(hide_grouplabel = TRUE) %>% 
          flextable_polish() %>% 
          flextable_autofit() %>% 
          padding(i = ~ label %in% c("Men", "Women"), 
                  j = 1,
                  padding.left = 15,
                  part = 'body') %>% 
          set_header_labels(
            label = "Variable",
            tbv_est = paste("Expected change (95% CI) in", lpa_lab_unit),
            tbv_pval = "P-value"
          ) %>% 
          add_footer_lines(
            values = as_paragraph(
              glue("Percentage of variance explained in \\
                   {lpa_lab_bare} by this model: {r_squared}")
            )
          )
        
      })
  ) %>% 
  transmute(
    name = paste(.lpa_defn, .lpa_trafo, .risk_horizon, sep = '_'),
    tbl
  ) %>% 
  deframe()


```

```{r create-tbl_lpa_cph}

tbl_lpa_cph <- analyses %>% 
  filter(.lpa_trafo == 'no_trafo') %>% 
  select(-predictions, -starts_with('fit_lm')) %>% 
  mutate(
    data_tbl = pmap(
      .l = list(.risk_type, .lpa_defn, .lpa_trafo, fit_cox_tidy),
      .f = function(.risk_type, .lpa_defn, .lpa_trafo, fit_cox_tidy){ 
        
        fit_cox_tidy %>% 
          filter(level != "no" | is.na(level)) %>% 
          select(-level, -term, -std.error, -statistic) %>% 
          mutate(name = translate_names(name, units = 'model'),
                 label = if_else(is.na(label) | label == 'yes', name, label),
                 name = if_else(name == label, NA_character_, name),
                 tbv_pval = table_pvalue(p.value),
                 tbv_est = table_glue("{estimate}\n({conf.low}, {conf.high})",
                                      rspec = rspec_2d),
                 tbv_est = if_else(reference, true = "1 (ref)", false = tbv_est),
                 .keep = 'unused') %>% 
          relocate(tbv_pval, .after = tbv_est) 
        
      } 
      
    )
  ) %>% 
  select(.risk_horizon, .risk_type, data_tbl) %>% 
  unnest(data_tbl) %>% 
  pivot_wider(names_from = .risk_type, 
              values_from = c(tbv_est, tbv_pval),
              names_vary = 'slowest') %>% 
  mutate(.risk_horizon = paste("year", .risk_horizon, sep = '_')) %>% 
  split(.$.risk_horizon) %>% 
  map(
    ~ .x %>% 
      select(-.risk_horizon) %>% 
      as_grouped_data(groups = 'name') %>% 
      drop_empty_rows() %>%  
      as_flextable(hide_grouplabel = TRUE) %>% 
      flextable_polish() %>% 
      flextable_autofit() %>% 
      padding(i = ~ label %in% c("Men", "Women"), 
              j = 1,
              padding.left = 15,
              part = 'body') %>% 
      set_header_labels(
        label = "Variable",
        tbv_est_ascvd = "Hazard ratio\n(95% CI)",
        tbv_pval_ascvd = "P-value",
        tbv_est_cvd = "Hazard ratio\n(95% CI)",
        tbv_pval_cvd = "P-value"
      ) %>% 
      add_header_row(values = c("Variable",
                                "ASCVD",
                                "CVD"),
                     colwidths = c(1, 2, 2)) %>% 
      merge_v(part = 'header')
  ) 



```


```{r create-tbl_cstats}

tbl_cstats <- evaluations %>% 
  mutate(.risk_horizon = paste("year", .risk_horizon, sep = '_')) %>% 
  split(.$.risk_horizon) %>% 
  map(
    ~ select(.x, -predictions) %>% 
      unnest(evaluation) %>% 
      filter(str_detect(stat, '^auc'),
             .lpa_defn == 'lpa_nmol',
             .lpa_trafo == 'no_trafo') %>% 
      mutate(tbv = table_glue("{est}\n({lwr}, {upr})", 
                              rspec = rspec_4d), 
             .keep = 'unused') %>% 
      select(group_level, .risk_type, model, stat, tbv) %>% 
      pivot_wider(names_from = model, values_from = tbv) %>% 
      mutate(group_level = translate_categories(group_level),
             .risk_type = str_to_upper(.risk_type),
             stat = recode(stat, 
                           "auc" = "C-statistic (95% CI)",
                           "auc_vs_prevent" = "Difference\n(95% CI)"),
             prevent = if_else(is.na(prevent), "0 (reference)", prevent)) %>% 
      relocate(.risk_type, .after = enhance) %>% 
      as_grouped_data(groups = '.risk_type') %>% 
      as_flextable(hide_grouplabel = TRUE) %>% 
      flextable_polish() %>% 
      flextable_autofit() %>% 
      set_header_labels(group_level = "Group",
                        stat = "Statistic",
                        prevent = "Base",
                        patch = "Patched with Lp(a)",
                        enhance = "Enhanced with Lp(a)") %>% 
      add_header_row(values = c("Group", "Statistic", "PREVENT Equations"),
                     colwidths = c(1, 1, 3)) %>% 
      merge_v(part = 'header') %>% 
      merge_v(part = 'body', j = 'group_level')
  )

```


```{r create-tbl_nri}

tbl_nri <- evaluations %>% 
  mutate(.risk_horizon = paste("year", .risk_horizon, sep = '_')) %>% 
  split(.$.risk_horizon) %>% 
  map(
    ~ select(.x, -predictions) %>% 
      unnest(evaluation) %>% 
      filter(str_detect(stat, '^nri'),
             .lpa_defn == 'lpa_nmol',
             .lpa_trafo == 'no_trafo') %>% 
      mutate(tbv = table_glue("{est}\n({lwr}, {upr})", 
                              rspec = rspec_3d), 
             .keep = 'unused') %>% 
      select(group_level, .risk_type, model, stat, tbv) %>% 
      pivot_wider(names_from = model, values_from = tbv) %>% 
      mutate(group_level = translate_categories(group_level),
             .risk_type = str_to_upper(.risk_type),
             stat = recode(stat, 
                           "nri_overall" = "NRI",
                           "nri_plus" = "NRI+",
                           "nri_minus" = "NRI-")) %>% 
      as_grouped_data(groups = '.risk_type') %>% 
      as_flextable(hide_grouplabel = TRUE) %>% 
      flextable_polish() %>% 
      flextable_autofit() %>% 
      set_header_labels(group_level = "Group",
                        stat = "Statistic",
                        prevent = "Base",
                        patch = "Patched with Lp(a)",
                        enhance = "Enhanced with Lp(a)") %>% 
      merge_v(part = 'header') %>% 
      merge_v(part = 'body', j = 'group_level') %>% 
      footnote(j = 'stat', i = 1, 
               part = 'header', 
               ref_symbols = "1", 
               value = footer_nri)
  )

```


```{r create-tbl_calib}

tbl_calib <- evaluations %>% 
  mutate(.risk_horizon = paste("year", .risk_horizon, sep = '_')) %>% 
  split(.$.risk_horizon) %>% 
  map(
    ~ select(.x, -predictions) %>% 
      unnest(evaluation) %>% 
      filter(str_detect(stat, '^calib'),
             .lpa_defn == 'lpa_nmol',
             .lpa_trafo == 'no_trafo') %>% 
      mutate(tbv = table_glue("{est}\n({lwr}, {upr})", 
                              rspec = rspec_3d), 
             .keep = 'unused') %>% 
      select(group_level, .risk_type, model, stat, tbv) %>% 
      pivot_wider(names_from = model, values_from = tbv) %>% 
      relocate(prevent, patch, enhance, .after = stat) %>% 
      mutate(group_level = translate_categories(group_level),
             .risk_type = str_to_upper(.risk_type),
             stat = recode(stat, 
                           "calib_int" = "Calibration intercept",
                           "calib_slp" = "Calibration slope")) %>% 
      as_grouped_data(groups = '.risk_type') %>% 
      as_flextable(hide_grouplabel = TRUE) %>% 
      flextable_polish() %>% 
      flextable_autofit() %>% 
      set_header_labels(group_level = "Group",
                        stat = "Statistic",
                        prevent = "Base",
                        patch = "Patched with Lp(a)",
                        enhance = "Enhanced with Lp(a)") %>% 
      merge_v(part = 'header') %>% 
      merge_v(part = 'body', j = 'group_level') 
  )

```


\newpage

# Primary objective 1: Apply the predictor patch method to incorporate molar concentration of Lp(a) into the 30-year PREVENT total CVD and ASCVD risk prediction equations.

\newpage

## Table 1: Participants included in 30-year risk analysis.

```{r}

exclusions_30yr %>% 
  set_names(translate_categories(names(.), names = 'study')) %>% 
  flextable() %>% 
  set_header_labels(
    inclusion = "Inclusion criteria",
    .overall = "Overall"
  ) %>% 
  flextable_polish() %>% 
  flextable_autofit()

```

\newpage

## Table 2: Characteristics of participants included in 30-year risk analysis

```{r}

tbl_characteristics$year_27.5

```

\newpage

## Table 3: Multi-variable adjusted linear regression estimate (95% CI) for Lp(a) in nmol/L among MESA participants

```{r}

tbl_lpa_lm$lpa_nmol_no_trafo_27.5

```

\newpage

## Table 4: Multi-variable adjusted hazard ratios for 30-year total CVD and ASCVD among MESA participants.

```{r}

tbl_lpa_cph$year_27.5

```

`r page_long_above()`

## Figure 1: Predicted 30-year risk with base PREVENT equations and patched PREVENT equations among ARIC participants

```{r}

data_gg <- evaluations %>% 
  filter(group_variable == 'overall',
         .risk_horizon > 10) %>% 
  select(-evaluation) %>% 
  unnest(predictions) %>% 
  filter(.lpa_defn == 'lpa_nmol', .lpa_trafo == 'no_trafo') %>% 
  select(.risk_type, prevent, patch) %>% 
  mutate(
    .risk_type = paste("Predicted 30-year risk of",
                       str_to_upper(.risk_type)),
    category = case_when(
      prevent > 0.20 & patch < 0.20 ~ "From >20% to < 20% w/patch",
      prevent > 0.20 & patch > 0.20 ~ ">20% in both",
      prevent < 0.20 & patch > 0.20 ~ "From <20% to >20% w/patch",
      prevent < 0.20 & patch < 0.20 ~ "<20% in both"
    )
  )

ggplot(data_gg) +
  aes(x = prevent, y = patch, color = category) + 
  geom_point() + 
  facet_wrap(~.risk_type) + 
  labs(x = "PREVENT risk score, %",
       y = "Patched PREVENT risk score, %",
       color = "Risk category") + 
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(labels = scales::percent) +
  theme_bw() + 
  theme(panel.grid = element_blank(),
        legend.position = 'top')

```

`r page_wide_above()`

# Primary objective 2: Compare the performance of the Lp(a) patched 30-year PREVENT total CVD and ASCVD risk prediction equations versus the original PREVENT equations and the approach of including molar concentration of Lp(a) as a risk-enhancing factor, as proposed in the 2022 AHA Lp(a) Scientific Statement.

\newpage

## Table 5: C-statistics of the PREVENT equations for 30-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.

```{r}
tbl_cstats$year_27.5
```

\newpage

## Table 6: Net reclassification improvement of the PREVENT equations for 30-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.


```{r}
tbl_nri$year_27.5
```

\newpage

## Table 7: Calibration slope and intercept of the PREVENT equations for 30-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.

```{r}
tbl_calib$year_27.5
```

\newpage
 
# Secondary objective 1: Apply the predictor patch method to incorporate molar concentration of Lp(a) into the 10-year PREVENT total CVD and ASCVD risk prediction equations. 

\newpage

## Table 8: Participants included in 10-year risk analysis.

```{r}

exclusions_10yr %>% 
  set_names(translate_categories(names(.), names = 'study')) %>% 
  flextable() %>% 
  set_header_labels(
    inclusion = "Inclusion criteria",
    .overall = "Overall"
  ) %>% 
  flextable_polish() %>% 
  flextable_autofit()

```


\newpage

## Table 9: Characteristics of participants included in 10-year risk analysis

```{r}

tbl_characteristics$year_10

```

\newpage

## Table 10: Multi-variable adjusted linear regression estimate (95% CI) for Lp(a) in nmol/L among MESA participants

```{r}

tbl_lpa_lm$lpa_nmol_no_trafo_10

```

\newpage

## Table 11: Multi-variable adjusted hazard ratios for 10-year total CVD and ASCVD among MESA participants.

```{r}

tbl_lpa_cph$year_10

```

`r page_long_above()`

## Figure 1: Predicted 10-year risk with base PREVENT equations and patched PREVENT equations among ARIC participants.

```{r}

data_gg <- evaluations %>% 
  filter(group_variable == 'overall',
         .risk_horizon == 10) %>% 
  select(-evaluation) %>% 
  unnest(predictions) %>% 
  filter(.lpa_defn == 'lpa_nmol', .lpa_trafo == 'no_trafo') %>% 
  select(.risk_type, prevent, patch) %>% 
  mutate(
    .risk_type = paste("Predicted 10-year risk of",
                       str_to_upper(.risk_type)),
    category = case_when(
      prevent > 0.10 & patch < 0.10 ~ "From >10% to < 10% w/patch",
      prevent > 0.10 & patch > 0.10 ~ ">10% in both",
      prevent < 0.10 & patch > 0.10 ~ "From <10% to >10% w/patch",
      prevent < 0.10 & patch < 0.10 ~ "<10% in both"
    )
  )

ggplot(data_gg) +
  aes(x = prevent, y = patch, color = category) + 
  geom_point() + 
  facet_wrap(~.risk_type) + 
  labs(x = "PREVENT risk score, %",
       y = "Patched PREVENT risk score, %",
       color = "Risk category") + 
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(labels = scales::percent) +
  theme_bw() + 
  theme(panel.grid = element_blank(),
        legend.position = 'top')

```

`r page_wide_above()`

# Secondary objective 2: Compare the performance of the Lp(a) patched 10-year PREVENT total CVD and ASCVD risk prediction equations versus the original PREVENT equations and the approach of including molar concentration of Lp(a) as a risk-enhancing factor, as proposed in the 2022 AHA Lp(a) Scientific Statement.  

\newpage

## Table 5: C-statistics of the PREVENT equations for 10-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.

```{r}
tbl_cstats$year_10
```

\newpage

## Table 6: Net reclassification improvement of the PREVENT equations for 10-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.


```{r}
tbl_nri$year_10
```

\newpage

## Table 7: Calibration slope and intercept of the PREVENT equations for 10-year risk prediction using the base equation and two modified equations that incorporate Lp(a) among ARIC participants.

```{r}
tbl_calib$year_10
```

\newpage

# Exploratory objective: Repeat the primary and secondary objectives using Lp(a) in mass concentration.


\newpage

# Appendix

`r page_long_above()`

## Figure A1: Cumulative distribution of numerical variables by study.

```{r}

data_gg <- data_train %>% 
  bind_rows(data_test) %>% 
  select(study, 
         where(is.numeric),
         -risk_horizon,
         -contains("hba1"),
         -matches("^time|^status")) %>% 
  pivot_longer(cols = -study) %>% 
  mutate(name = translate_names(name),
         study = translate_categories(study, name = 'study'))

ggplot(data_gg, aes(x = value, color = study)) + 
  stat_ecdf(geom = 'step') + 
  facet_wrap(~name, scales = 'free_x') + 
  theme(legend.position = 'top')


```

\newpage

## Figure A2: Box-plots of numerical variables by study.

```{r}

ggplot(data_gg, aes(x = study, y = value, fill = study)) + 
  geom_boxplot() + 
  facet_wrap(~name, scales = 'free_y')  + 
  theme(legend.position = 'top')


```

\newpage

## Figure A3: Expected CVD risk by Lp(a) using natural splines

```{r}

data_gg <- fig_lpa_spline %>% 
  mutate(data_type = recode(data_type, 
                            "test" = "Testing",
                            "train" = "Training"),
         risk_horizon = recode(risk_horizon,
                               '10' = '10-year risk',
                               '27.5' = '30-year risk'))

ggplot(data_gg) +
  aes(x = lpa_nmol, y = pred, color = data_type) +
  geom_line() +
  facet_wrap(~risk_horizon) +
  labs(x = "Lp(a), nMol/L",
       y = "CVD risk, %",
       color = "Data") + 
  scale_y_continuous(labels = scales::percent)

```


`r page_wide_above()`
